# Generating data using gcipdr methods

In this chapter we will look at how to generate synthetic data on the client side using DataSHIELD functions to extract characteristics of the data set from the server side.

The **gcipdr** package is an initial version and more details can be found [here](https://github.com/bonorico/gcipdr).

More prerequisites:

```{r install gcipdr}

library(devtools)
install_github("bonorico/gcipdr")

url <- "https://cran.r-project.org/src/contrib/Archive/JohnsonDistribution/JohnsonDistribution_0.24.tar.gz"
pkgFile <- "JohnsonDistribution_0.24.tar.gz"
download.file(url = url, destfile = pkgFile)

install.packages(pkgs=pkgFile, type="source", repos=NULL)

unlink(pkgFile)

```
Again we need to build a login object for the server that holds the data:

```{r log in object gcipdr}
builder <- DSI::newDSLoginBuilder()
# hide credentials
builder$append(server="server1", url="https://opal-sandbox.mrc-epid.cam.ac.uk",
               user="dsuser", password="password", 
               table = "DASIM.DASIM1")
logindata <- builder$build()
```

And then we establish a connection to the server:

```{r log in gcipdr}
library(DSOpal)
if(exists("connections")){
  datashield.logout(conns = connections)
}
connections <- datashield.login(logins=logindata, assign = TRUE)
```

