# Generating data using simstudy methods {#simstudy}

In this chapter we will look at how to generate synthetic data on the client side using DataSHIELD functions to extract summary characteristics of the data set from the server side.

## Overview

The steps for using the **simstudy** to generate a synthetic data set:

1. Connect to the server
2. Split factor variables in dummy (binary) variables
3. Prepare lists of continuous and factor variables
4. Generate synthetic data set
5. Recombine dummy variables.

## Details

Again we need to build a login object for the server that holds the data. Note that the **dsSynthetic** functions have been written to work with a single connection:

```{r log in object simstudy}
builder <- DSI::newDSLoginBuilder()
# hide credentials
builder$append(server="server1", url="https://opal-sandbox.mrc-epid.cam.ac.uk",
               user="dsuser", password="P@ssw0rd", 
               table = "DASIM.DASIM1")
logindata <- builder$build()
```

And then we establish a connection to the server:

```{r message=FALSE, "log in simstudy"}
library(DSOpal)
if(exists("connections")){
  datashield.logout(conns = connections)
}
connections <- datashield.login(logins=logindata, assign = TRUE)
```
This method builds on the fundamentals of the **simstudy** package. There is a good introduction to this package [here](https://kgoldfeld.github.io/simstudy/index.html).

First we load the library
```{r simstudy}
library("simstudy")
library("dsBaseClient")
library("dsSyntheticClient")
```

The **simstudy** method uses a function called `genCorFlex` (see **simstudy** help for more details). This function is currently not able to support categorical distributions, but does support binary distributions. Therefore we need to convert factor variables with more than 2 levels into a series of binary dummy variables. For example, the `PM_BMI_CATEGORICAL` variable has levels 1, 2 and 3. This will need to be converted into 2 binary variables with 2 levels. A helper function is provided to make this process easier.

```{r message=FALSE, "prep binaries"}
ds.binary.helper(dataframe = "D", factor_variables = "PM_BMI_CATEGORICAL", newobj = "my.bin", datasources = connections)
```
Looking at the new dataframe, which has the 2 binary variables in it:

```{r message=FALSE}
ds.summary("my.bin", datasources = connections)
```
We will make a new dataframe without the original `PM_BMI_CATEGORICAL` variable but with the new binaries:
```{r warning=FALSE, message=FALSE}
vars = c("LAB_TSC", "LAB_TRIG", "LAB_HDL", "LAB_GLUC_FASTING", "PM_BMI_CONTINUOUS", "DIS_CVA", "DIS_DIAB", "DIS_AMI", "GENDER")
ds.subset(x="D", subset="D2", cols=vars, datasources = connections)
ds.dataFrame(x = c("D2","my.bin"), newobj = "new_frame", datasources = connections)
```
Now we are ready to generate the synthetic data using the `ds.genCorFlex.helper` function. We need to provide a vector of continuous variables and factor variables:
```{r warning=FALSE,message=FALSE}
cont_vars = c("LAB_TSC", "LAB_TRIG", "LAB_HDL", "LAB_GLUC_FASTING", "PM_BMI_CONTINUOUS")
factor_vars = c( "DIS_CVA", "DIS_DIAB", "DIS_AMI", "GENDER", "X1_PM_BMI_CATEGORICAL", "X2_PM_BMI_CATEGORICAL")
dd = ds.genCorFlex.helper(dataframe = "new_frame", cont_variables = cont_vars, factor_variables = factor_vars, num_rows = 10000, datasources = connections)
```
If all has gone well the synthetic data should have a similar structure and properties to the real data:
```{r}
head(dd)

```
We can reconstruct the `PM_BMI_CATEGORICAL` variable as follows:
```{r}
dd$PM_BMI_CATEGORICAL = 1 + dd$X1_PM_BMI_CATEGORICAL_num + dd$X2_PM_BMI_CATEGORICAL_num
```
Such that it now has levels 1, 2 and 3.

## Brief comments on the validity of the data

In the example above we chose to synthetically generate the variable `PM_BMI_CATEGORICAL`. This variable is actually derived from the continuous variable `PM_BMI_CONTINUOUS`, with (say) BMI <25 being category 1, 25<= BMI < 30 being category 2 etc. Because of the probabilistic way in which the data are generated, this categorisation is not enforced in the data synthesis. It might be better to generate the continuous variable only and add the categorical variable afterwards.

## Comparing **synthpop** and **simstudy** approaches

The main difference between the **synthpop** and **simstudy** approaches is that the synthesis takes place on the server side with **synthpop** and on the client side with **simstudy**.

TO DO: Computational load, what algorithms, amount of control
